---
title: "Example PSRC map"
author: "suzanne"
date: "2023-05-05"
output: html_document
---
# How to quickly make a formatted map
Inputs required: 
a data table with a field to map and an ID, 
a geographic layer with a matching ID
User options:
Map title, Map subtitle,
Legend tile, Legend subtitle
psrccolors



```{r }
library(sf)
library(dplyr)
library(psrccensus)
library(psrcelmer)
library(psrcplot)
Sys.getenv("CENSUS_API_KEY")
```


This function creates a leaflet map for you.
```{r}
create_psrc_map <- function(lyr,
                            lyr_data_field,
                            map_title, map_subtitle,
                            legend_title, legend_subtitle,
                            psrc_col_pal,
                            map_title_position='topleft',
                            map_lat=47.615, map_lon=-122.257, 
                            map_zoom=8.5, wgs84=4326){

  # psrc colors need more contrast to work
  #pal <- leaflet::colorNumeric(palette=psrc_col_pal, domain = lyr_datafield)
  
   pal<-leaflet::colorNumeric(palette="Blues", domain = lyr_datafield)
  
  labels <- paste0('Estimate: ', prettyNum(round(lyr_datafield, -1), big.mark = ",")) %>%
    lapply(htmltools::HTML)
  
  m <- leaflet::leaflet() %>%
    leaflet::addMapPane(name = "polygons", zIndex = 410) %>%
    leaflet::addMapPane(name = "maplabels", zIndex = 500) %>% # higher zIndex rendered on top
    
    leaflet::addProviderTiles("CartoDB.VoyagerNoLabels") %>%
    leaflet::addProviderTiles("CartoDB.VoyagerOnlyLabels",
                              options = leaflet::leafletOptions(pane = "maplabels"),
                              group = "Labels") %>%
    
    leaflet::addEasyButton(leaflet::easyButton(icon="fa-globe",
                                               title="Region",
                                               onClick=leaflet::JS("function(btn, map){map.setView([47.615,-122.257],8.5); }"))) %>%
    leaflet::addPolygons(data=lyr,
                         fillOpacity = 0.7,
                         fillColor = pal(lyr_datafield),
                         weight = 0.7,
                         color = "#BCBEC0",
                         group="estimate",
                         opacity = 0,
                         stroke=FALSE,
                         options = leaflet::leafletOptions(pane = "polygons"),
                         dashArray = "",
                         highlight = leaflet::highlightOptions(
                           weight =5,
                           color = "76787A",
                           dashArray ="",
                           fillOpacity = 0.7,
                           bringToFront = TRUE),
                         label = labels,
                         labelOptions = leaflet::labelOptions(
                           style = list("font-weight" = "normal", padding = "3px 8px"),
                           textsize = "15px",
                           direction = "auto")) %>%
    
     leaflet::addLegend(pal = pal,
                        values = lyr_datafield,
                        position = "bottomright",
                        title = paste(legend_title, '<br>', legend_subtitle)) %>%
     
    leaflet::addControl(html = paste(map_title, '<br>', map_subtitle),
                        position = map_title_position,
                        layerId = 'mapTitle') %>%
    
    leaflet::addLayersControl(baseGroups = "CartoDB.VoyagerNoLabels",
                              overlayGroups = c("Labels", "estimate")) %>%
    
    leaflet::setView(lng=map_lon, lat=map_lat, zoom=map_zoom)
  
  return(m)
  
}
```

## Step One: Read in Data Table with a geographic field ID
Prepare data to join with the geographic layer
Note: you will need field names to match on


### Get the data from Census
```{r pressure, echo=FALSE}
big_tbl <-psrccensus::get_acs_recs(geography='tract',table.names='B02015',year=2021, acs.type='acs5')
```

### Clean up the data to join to the geographic layer
```{r, echo=FALSE}
# read in data table and process data to join and map
tbl <- big_tbl %>%filter(label=='Estimate!!Total:!!Korean')%>% 
  dplyr::select(GEOID,estimate) %>%
  dplyr::mutate(dplyr::across(c('GEOID'), as.character))%>%
  dplyr::group_by(GEOID) %>%
  dplyr::summarise(Total=sum(estimate))
```


##Read ElmerGeo geographic layer

```{r pressure, echo=FALSE}
tract_layer_name <- "TRACT2020_NOWATER"

lyr <- st_read_elmergeo(tract_layer_name)
```

Join to data table layer via matching IDs, choose the field to map
```{r pressure, echo=FALSE}
lyr_data<- dplyr::left_join(lyr,tbl, by = c("geoid20"="GEOID"))
# this is the field to map
lyr_datafield<-lyr_data$Total
```

```{r}
map_it<-create_psrc_map(lyr=lyr_data, 
                        lyr_data_field=lyr_data_field,
                        map_title= 'Korean Population',
                        map_subtitle = 'ACS 2017-2021',
                        legend_title= 'Korean Population',
                        legend_subtitle = 'by Census Tract',
                        psrc_col_pal=psrc_colors$blues_inc)

```

```{r}
map_it
```


